---
output: github_document
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


# penguinDash

`penguinDash` is an R package built for the Monash University unit ETC5523 – Communicating with Data.

It provides an interactive dashboard for exploring the well-known `palmerpenguins` dataset, which contains data for three penguin species observed in the Palmer Archipelago, Antarctica.

### What it does

The package helps users visualize and understand the relationships between various physical measurements of penguins.

It includes:

*   A cleaned dataset, `penguins_clean`, derived from the `palmerpenguins` R package.
*   An interactive dashboard launched by the function `launch_dashboard()`.
*   A vignette that provides a guide to the dataset and how to use the dashboard effectively.

### Installation and Use

#### Install and load

You can install `penguinDash` from its GitHub repository and load it into your R session.
```{r, eval=FALSE}
install.packages("remotes")
remotes::install_github("ETC5523-2025/assignment-4-packages-and-shiny-apps-SlothGodCh")

library(penguinDash)
```

#### Access the data

The cleaned dataset is included with the package and can be accessed directly after loading.

```{r,eval=FALSE}
# Preview the first few rows of the dataset
head(penguins_clean)
```

#### Launch the dashboard

The main feature of the package is its interactive Shiny dashboard.

```{r,eval=FALSE}
# This command will open the dashboard in your default web browser
launch_dashboard()
```

### Data Source

The data was collected and made available by Dr. Kristen Gorman, Dr. Allison Horst, and Dr. Alison Hill, and is available in the R package `palmerpenguins`.

Horst AM, Hill AP, Gorman KB (2020). *palmerpenguins: Palmer Archipelago (Antarctica) penguin data*. R package version 0.1.0. <https://allisonhorst.github.io/palmerpenguins/>

### Interpretation Example

Using the dashboard reveals distinct patterns among the penguin species. For example, by plotting flipper length against body mass, you can clearly see that Gentoo penguins are generally larger and have longer flippers than both Adelie and Chinstrap penguins. The interactive filters allow for a deeper exploration of these relationships, such as observing how bill dimensions differ between male and female penguins within the same species.

### License

MIT + file LICENSE
Author: Jesus Biurrun

## Website

The pkgdown website for this package is available at:
https://etc5523-2025.github.io/assignment-4-packages-and-shiny-apps-SlothGodCh/


