---
title: "Introduction to penguinDash"
author: "Jesus Biurrun"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to penguinDash}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
### Introduction

The `penguinDash` package provides an interactive Shiny dashboard for exploring the `palmerpenguins` dataset. This vignette serves as a comprehensive guide to installing the package, loading the included data, and launching the interactive application.

### Installation

The development version of `penguinDash` can be installed directly from its GitHub repository. You will need the `remotes` package to do this.
```{r, eval=FALSE}
install.packages("remotes")
remotes::install_github("ETC5523-2025/assignment-4-packages-and-shiny-apps-SlothGodCh")
```
### Using the Package

#### Loading the Package and Data

Once installed, load the package into your R session to get started.

```{r, eval=FALSE}
library(penguinDash)
```
#### Launching the Interactive Dashboard

The main feature of `penguinDash` is its interactive dashboard. To launch it, simply run the `launch_dashboard()` function.
```{r, eval=FALSE}
# This command will open the Shiny app in a new window or your browser
launch_dashboard()
```



### A Tour of the Dashboard

The **Palmer Penguins Explorer** dashboard provides several interactive controls to help you explore the data:

* **Dark Mode Toggle**: Click the sun/moon icon in the top-right corner to switch between light and dark themes.

* **Species Filter**: A dropdown menu allows you to focus on a single species (Adelie, Chinstrap, or Gentoo) or view all species together.

* **X-Axis Variable**: Choose which measurement to display on the horizontal axis (Bill Length, Bill Depth, Flipper Length, or Body Mass).

* **Y-Axis Variable**: Choose which measurement to display on the vertical axis, enabling comparison of any two variables.

* **Color by Species**: A checkbox that toggles species coloring on or off, useful for viewing overall patterns or species-specific differences.

* **Bill Length Filter**: A slider allows you to narrow the data to a specific range of bill lengths.

The dashboard features three main sections:

1. **KPI Cards**: Display summary statistics (average body mass, average flipper length, and number of penguins shown) that update based on your filters.

2. **Interactive Plot**: A dynamic scatter plot with hover tooltips showing detailed information for each penguin.

3. **Data Table**: A filterable, sortable table showing all penguins that match your current selection.

### Example: Finding Insights

The dashboard makes it easy to discover patterns. For instance:

* By selecting **Gentoo** species and comparing **Flipper Length** vs **Body Mass**, you can observe a strong positive correlation, confirming that Gentoo penguins with longer flippers tend to have higher body mass.

* By toggling **Color by Species** off, you can examine overall trends across all penguins without species distinction.

* Using **dark mode** enhances viewing comfort during extended analysis sessions.

Comparing different species reveals distinct clusters and differences in size and proportions across the Palmer Archipelago penguin populations.

### Data Source

This package uses data from the `palmerpenguins` R package.

*Horst AM, Hill AP, Gorman KB (2020). palmerpenguins: Palmer Archipelago (Antarctica) penguin data. R package version 0.1.0. https://allisonhorst.github.io/palmerpenguins/*

### Additional Resources

For more information about the package, visit the pkgdown website at:
https://etc5523-2025.github.io/assignment-4-packages-and-shiny-apps-SlothGodCh/
