---
title: "Introduction to penguinDash"
author: "Jesus Biurrun"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to penguinDash}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
### Introduction

The `penguinDash` package provides an interactive Shiny dashboard for exploring the `palmerpenguins` dataset. This vignette serves as a comprehensive guide to installing the package, loading the included data, and launching the interactive application.

### Installation

The development version of `penguinDash` can be installed directly from its GitHub repository. You will need the `remotes` package to do this.
```{r, eval=FALSE}
install.packages("remotes")
remotes::install_github("ETC5523-2025/assignment-4-packages-and-shiny-apps-SlothGodCh")
```
### Using the Package

#### Loading the Package and Data

Once installed, load the package into your R session to get started.

```{r, eval=FALSE}
library(penguinDash)
```
#### Launching the Interactive Dashboard

The main feature of `penguinDash` is its interactive dashboard. To launch it, simply run the `launch_dashboard()` function.
```{r, eval=FALSE}
# This command will open the Shiny app in a new window or your browser
launch_dashboard()
```


### A Tour of the Dashboard

The **Palmer Penguins Explorer** dashboard provides several interactive controls to help you explore the data:

*   **Filter by Bill Length**: A slider allows you to focus on a specific range of bill lengths.
*   **Select Species**: Checkboxes let you include or exclude the Adelie, Chinstrap, and Gentoo species from the visualizations.
*   **Select Y-Axis Variable**: A dropdown menu allows you to change the variable on the plot's y-axis, enabling you to compare bill length against bill depth, flipper length, or body mass.

The dashboard's main panel displays a reactive scatter plot and a data table that update instantly based on your selections.

### Example: Finding Insights

The dashboard makes it easy to discover patterns. For instance, by selecting only the "Gentoo" species and plotting "flipper_length_mm" against "body_mass_g", you can observe a strong positive correlation, confirming that Gentoo penguins with longer flippers tend to have a higher body mass. Comparing this to the other species reveals distinct clusters and differences in size and proportions.

### Data Source

This package uses data from the `palmerpenguins` R package.

*Horst AM, Hill AP, Gorman KB (2020). palmerpenguins: Palmer Archipelago (Antarctica) penguin data. R package version 0.1.0.*
<https://allisonhorst.github.io/palmerpenguins/>









